### 2021-10-21 text/outline/outline.tex
proof of monotonicity in beta

\documentclass[12pt]{article}
\usepackage{amsfonts}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{xcolor}
\newcommand{\blue}[1]{\textcolor{blue}{#1}}
\usepackage{float}
\usepackage[caption = false]{subfig}
\usepackage{/Users/timbarry/optionFiles/mymacros}
\DeclareMathOperator*{\argmin}{arg\,min}

\begin{document}
\noindent
Tim, Gene, Kathryn
\begin{center}
\textbf{CRISPR genome editing, single-cell sequencing, and exponential family measurement error models}
\end{center}

@@

Our work makes two main contributions. First, we study the thresholding method from empirical and theoretical perspectives. Notably, we demonstrate on real data that the thresholding method exhibits attenuation bias and a bias-variance tradeoff as a function of the selected threshold, and we recover these phenomena in precise mathematical terms in an idealized Gaussian model. Second, we introduce a new method for estimation and inference in single-cell CRISPR screens that accounts for the measurement error inherent in the experiment. The method, called \textit{GLM-EIV} (generalized linear model with errors-in-variables), implicitly estimates the probability that each cell was perturbed, obviating the need to explicitly impute perturbation assignments via thresholding or another heuristic. Theoretical analyses and simulation studies indicate that GLM-EIV outperforms the thresholding method in large regions of the parameter space.

@@

\begin{equation}\label{theoretical_model}
\begin{cases}
m_i = \beta^m_0 + \beta^m_1 p_i + \ep_i \\
g_i = \beta^g_0 + \beta^g_1 p_i + \tau_i \\
p_i \sim \textrm{Bern}(\pi) \\
\ep_i, \tau_i \sim N(0,1) \\
p_i \indep \tau_i \indep \ep_i.
\end{cases}
\end{equation}
For a given threshold $c \in \R$, let $\hat{p}_i = \mathbb{I}(g_i \geq c).$ The thresholding estimator $\hat{\beta}^m_1$ for $\beta^m_1$ is given by $$\hat{\beta}^m_1 = \frac{\sum_{i=1}^n (\hat{p}_i - \overline{\hat{p}}) (m_i - \overline{m})}{\sum_{i=1}^n (\hat{p}_i - \overline{\hat{p}})}.$$ The limit in probability (as $n \to \infty$) of $\hat{\beta}^m_0$ is
\begin{equation}\label{thresh_est_intercepts}
\hat{\beta}^m_1 \xrightarrow{P} \beta^m_1 \left(\frac{ \pi( \omega - \E[ \hat{p}_i ])}{ \E[\hat{p}_i] (1 - \E[\hat{p}_i])}\right),
\end{equation}
where
$$\begin{cases}
\omega = 1 - \Phi\left(c - \beta_1^g - \beta_0^g \right) ,\\ \zeta = 1 - \Phi\left( c - \beta^g_0 \right), \\ \E[\hat{p}_i] = \zeta(1-\pi) + \omega\pi. 
\end{cases}$$
Let $\gamma: \R^4 \to \R$ be defined by
$$ \gamma( \beta^g_0, \beta^g_1, c, \pi ) = \frac{\pi (\omega - \E[\hat{p}_i])}{\E[\hat{p}_i] (1 - \E[\hat{p}_i]) }.$$ We call $\gamma$ the \textit{attenuation function}. Observe that \begin{itemize}
\item[i.] $\gamma$ does not depend on $\beta^m_1$ or $\beta^m_0$, and
\item[ii.] $\hat{\beta}^m_1 \xrightarrow{P} [\gamma(\beta_0^g, \beta_1^g, c, \pi)] \beta^m_1.$
\end{itemize}
Let $b: \R^4 \to \R$ be the asymptotic bias of $\hat{\beta}^m_1$. We can express $b$ in terms of $\gamma$ as $$b\left(\beta^g_0, \beta^g_1, c, \pi \right) = \beta^m_1 - \gamma( \beta_0^g, \beta_1^g, c, \pi) \beta^m_1 = \beta^m_1 \left[ 1- \gamma(\beta^g_0, \beta^g_1, c, \pi) \right].$$ The asymptotic bias vanishes when the attenuation function equals $1$.

@@

\begin{itemize}
\item[1.] The Bayes-optimal threshold $c_\textrm{bayes}$ is a critical value of the bias function $b$ (Figure \ref{thresholding_theoretical}a, all panels), i.e. 
$$ \frac{\partial b( \beta^g_0, \beta^g_1, c, 1/2)}{\partial c} \vline_{c = c_\textrm{bayes}} = 0.$$
\item[2.] For certain values of $\beta^g_0$ and $\beta^g_1$, $c_\textrm{bayes}$ is a maximizer of the bias (Figure \ref{thresholding_theoretical}a, left); for other values of $\beta^g_0$ and $\beta^g_1$, $c_\textrm{bayes}$ is a minimizer of the bias (Figure \ref{thresholding_theoretical}a, right).
\item[3.] The limit of the bias in $c$ is $1/2$, i.e. 
$$\lim_{c \to \infty} b(\beta^g_0, \beta^g_1, c, 1/2) = 1/2.$$ In other words, we always can set the threshold to a large number and achieve a bias of $1/2$ (Figure \ref{thresholding_theoretical}a, all panels). This fact establishes an upper bound on the bias of thresholded regression (under optimal threshold selection).
\item[4.] For $\beta^g_1 < 2 \Phi(3/4) \approx 1.35$, $c = c_\textrm{bayes}$ yields a \textit{larger} bias than $c = \infty$ (Figure \ref{thresholding_theoretical}a, left); for $\beta^g_1 > 2 \Phi(3/4)$, $c_\textrm{bayes}$ yields a \textit{smaller} bias than $c = \infty$ (Figure \ref{thresholding_theoretical}a, right); finally, for $\beta^g_1 = 2 \Phi(3/4)$, $c = c_\textrm{bayes}$ and $c = \infty$ both yield a bias of $1/2$ (Figure \ref{thresholding_theoretical}a, middle).
\item[5.] The bias decreases monotonically as $\beta_1^g$ increases (Figure \ref{thresholding_theoretical}a; fix threshold and scan panels from left to right). This is consistent with the intuition that the problem becomes easier as the gRNA mixture distribution becomes increasingly well-separated.
\item[6.] For all $\beta^g_0, \beta^g_1,$ and $c \in \R,$ the bias is strictly greater than $0$ and less than $2$ (Figure \ref{thresholding_theoretical}a, all panels). Therefore, the thresholding method suffers from attenuation bias.% a common phenomenon in errors-in-variables models.
\end{itemize}

@@

Next, we generalize the expression for bias when the threshold is large to arbitrary $\pi \in [0,1/2]$. The following simple relationship holds:
$$ \lim_{ c \to \infty } b(\beta^g_0, \beta^g_1, c, \pi) = \pi.$$ In other words, when the perturbation probability is equal to $\pi$, setting the threshold to a large number yields a bias of $\pi$ (Figure \ref{thresholding_theoretical}b). We can build intuition for this result by considering an extreme example. When $\pi$ is very small (e.g., $\pi = 0.01$), almost all cells are unperturbed. Therefore, in selecting a large threshold, one correctly classifies nearly all unperturbed cells as unperturbed; on the other hand, the \textit{perturbed} cells that one misclassifies as \textit{unperturbed} are swamped in number by the truly unperturbed cells, resulting in a small bias.

@@

Finally, to shed light on limitations of the large threshold selection strategy, we study the variance of the thresholding estimator. Let $v$ denote the limiting variance of scaled estimator $\sqrt{n} \hat{\beta}^m_1$, i.e.,
$$v(\beta^g_0, \beta^g_1, c, \pi) = \lim_{n \to \infty} \V(\sqrt{n} \hat{\beta}^m_1).$$ The limiting variance of $\sqrt{n} \hat{\beta}^m_1$ grows without bound as the threshold $c$ tends to infinity: 
$$ \lim_{c \to \infty} v(\beta^g_0, \beta^g_1, c, \pi) = \infty.$$ Thus, while selecting a large threshold is effective from a bias reduction standpoint when $\pi$ is small, it causes the variance to explode.

To more precisely characterize the bias-variance tradeoff of the thresholding method, we consider a slightly simpler model: we set $\beta^m_0$ and $\beta^g_0$ to zero in (\ref{theoretical_model}) and study the no-intercept least squares estimator
$$ \hat{\beta}^m_1 =  \frac{ \sum_{i=1}^n \hat{p}_i m_i }{ \sum_{i=1}^n \hat{p}_i^2}.$$ We uncover an exact bias-variance tradeoff as a function of the selected threshold (Figure \ref{thresholding_theoretical}c). The bias-variance analysis reveals that the best strategy for minimizing mean-squared error is to select a threshold that induces moderate bias. (A downside of this approach, of course, is that constructing valid confidence intervals becomes more challenging.)

@@

\subsubsection{Bayes-optimal decision boundary as a critical value of $\gamma$}
Let $c_\textrm{bayes} = \beta^g_0 + (1/2)\beta^g_1.$ We show that $c = c_\textrm{bayes}$ is a critical value of $\gamma$ for $\pi = 1/2$ and given $\beta^g_1$, i.e, $$D_2 \gamma (\beta^g_1, c_\textrm{bayes}, 1/2) = 0.$$ By the quotient rule,
\begin{multline}\label{quotient_rule}
D_2\gamma(\beta^g_1, c, 1/2) \\ = \frac{(1/2) D_2g(\beta^g_1, c, 1/2) h(\beta^g_1, c, 1/2) - (1/2) g(\beta^g_1, c, 1/2) D_2 h(\beta^g_1, c, 1/2)}{[h(\beta^g_1, c, \pi)]^2} \\ = \frac{\left[4(1/2) D_2g(\beta^g_1, c, 1/2)\right] h(\beta^g_1, c, 1/2) - (1/2) g(\beta^g_1, c, 1/2) \left[4 D_2 h(\beta^g_1, c, 1/2)\right]}{[2h(\beta^g_1, c, \pi)]^2},
\end{multline}
where the second equality multiplies both numerator and denominator by $4$. We have by (\ref{dg_dc}) that
\begin{multline}\label{dg_dc_bayes}
4 \cdot (1/2) \cdot D_2 g(\beta^g_1, c_\textrm{bayes}, 1/2)  = f( \beta^g_1/2) - f( -\beta^g_1/2) \\ = f(\beta^g_1/2) - f(\beta^g_1/2) = 0.
\end{multline}
Next, we have by (\ref{dh_dc}) that
\begin{multline}\label{dh_dc_bayes}
4 \cdot D_2 h(\beta^g_1, c_\textrm{bayes}, \pi) = [f( \beta^g_1/2) + f( -\beta^g_1/2)] \cdot \\ \left[  \int_{-\infty}^{-\beta^g_1/2} f + \int_{-\infty}^{\beta^g_1/2} f - \int_{-\infty}^{\beta^g_1/2} f - \int_{-\infty}^{-\beta^g_1/2} f \right] = 0.
\end{multline}
Plugging in (\ref{dh_dc_bayes}) and (\ref{dg_dc_bayes}) to (\ref{quotient_rule}), we find that 
$$D_2(\beta^g_1, c_\textrm{bayes}, 1/2) = 0.$$

% $$g'(\cdot, c) = - f( \beta^g_0 + \beta^g_1 - c ) + f(\beta^g_0 - c).$$ Second, 
% \begin{multline*}
% h'(\cdot, c) = -[f(\beta^g_0 - c)  + f( \beta^g_0 + \beta^g_1 - c )]  \left[\left( \int_{-\infty}^{c - \beta^g_0} f \right) + \left(\int_{-\infty}^{c - \beta_0^g - \beta_1^g} f \right) \right] \\ +[ f(c - \beta^g_0) + f(c - \beta^g_0 - \beta^g_1) ] \left[ \left( \int_{-\infty}^{ \beta_0^g - c} f \right) + \left( \int_{-\infty}^{ \beta^g_0 + \beta^g_1 - c} f \right) \right].
% \end{multline*}
% Evaluating $g'(\cdot, c)$ and $h'(\cdot, c)$ at $c = c_\textrm{bayes} $, we find that
% $$g'(\beta^g_0 + (1/2) \beta^g_1) = -f(\beta^g_1/2) + f(- \beta^g_1/2) = -f(\beta^g_1/2) + f(\beta^g_1/2) = 0,$$ where the second equality follows from the evenness of $f$. Similarly, 
% \begin{multline*}
% h'(\beta^g_0 + (1/2) \beta^g_1) = - \left[ f( -\beta^g_1/2) + f( \beta^g_1/2 ) \right] \left[ \int_{-\infty}^{ \beta^g_1/2} f + \int_{-\infty}^{ -\beta^g_1/2} f \right] + \\ \left[ f( \beta^g_1/2) + f( -\beta^g_1/2 ) \right] \left[ \int_{-\infty}^{ -\beta^g_1/2} f + \int_{-\infty}^{ \beta^g_1/2 f} \right] = 0.
% \end{multline*}
% We conclude that
% $$ \frac{\partial \gamma(\beta^g_0, \beta^g_1, c, 1/2)}{ \partial c} \vline_{[c = c_\textrm{bayes}]} = \frac{0\cdot h(\cdot, c_\textrm{bayes}) - g(\cdot, c_\textrm{bayes}) \cdot 0}{ h(\cdot, c_\textrm{bayes})^2} = 0.$$ Finally, because asymptotic bias $b$ can be expressed in terms of $\gamma$ as $$b(\beta^g_0. \beta^g_1, c, \pi) = \beta^m_1 \left[ 1 - \gamma(\beta^g_0, \beta^g_1, c, \pi) \right],$$ it follows that $c = c_\textrm{bayes}$ is a critical value of $b$ for $\pi = 1/2$ and fixed $(\beta^g_0, \beta^g_1)$.

% \subsubsection{Limit in $\beta^g_1$}

% We compute the limit of $\gamma(\beta^g_0, \beta^g_1, c, 1/2)$ in $\beta^g_1.$ We have that

% \begin{multline*}
 %\lim_{\beta^g_1 \to \infty } \gamma(\beta^g_0, \beta^g_1, c, 1/2) = \frac{ 1 - \int_{-\infty}^{\beta^g_0 - c } f}{\left[ 1 + \int_{-\infty}^{ \beta^g_0 - c} f \right] \left[\int_{-\infty}^{ c - \beta^g_0} f \right]} \\ = \frac{\int_{-\infty}^{c - \beta^g_0} f }{ \left[1 + \int_{-\infty}^{ \beta^g_0 - c} f \right] \left[\int_{-\infty}^{ c - \beta^g_0} f \right]} = \frac{1}{1 + \int_{-\infty}^{\beta^g_0 - c} f} < 1.
% \end{multline*}

% Similarly,
% \begin{multline*}
% \lim_{\beta^g_1 \to -\infty} \gamma(\beta^g_0, \beta^g_1, c, 1/2) = \frac{ -\int_{-\infty}^{\beta^g_0 - c} f }{\left[ \int_{-\infty}^{\beta^g_0 - c} f \right] \left[ \int_{-\infty}^{c - \beta^g_0} f + 1 \right]} = \frac{-1}{1 + \int_{-\infty}^{c - \beta^g_0} f} > -1.
% \end{multline*}

% \subsubsection{Monotonicity in $\beta^g_1$}
% We prove that $\gamma$ is monotonic in $\beta_1^g$ for given $(\beta^g_0, c)$ and $\pi = 1/2$. Recall that
% $$\gamma(\beta^g_0, \beta^g_1, c, 1/2) = \frac{ g(\beta^g_1, c) }{ h(\beta^g_1, c)}.$$ By the quotient rule,
% $$D_2 \gamma (\beta^g_0, \beta^g_1, c, 1/2) = \frac{D_1 g( \beta^g_1, c) h(\beta^g_1, c) - g(\beta^1_g, c) D_1 h(\beta^g_1, c) }{ [ h(\beta^g_1, c)^2]}.$$ The derivative $D_1 g(\beta^g_1, c)$ of $g$ is 
% $$D_1 g(\beta^g_1, c) = f( \beta^g_0 + \beta^g_1 - c ).$$

>>>>

\documentclass[12pt]{article}
\usepackage{amsfonts}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{xcolor}
\newtheorem{proposition}{Proposition}
\newtheorem{lemma}{Lemma}
\newcommand{\blue}[1]{\textcolor{blue}{#1}}
\usepackage{float}
\usepackage[caption = false]{subfig}
\usepackage{/Users/timbarry/optionFiles/mymacros}
\DeclareMathOperator*{\argmin}{arg\,min}

\begin{document}
\noindent
Tim, Gene, Kathryn
\begin{center}
\textbf{CRISPR genome engineering, single-cell sequencing, and exponential family measurement error models}
\end{center}

@@

Our work makes two main contributions. First, we conduct an in-depth study of the thresholding method from empirical and theoretical perspectives. Notably, we demonstrate on real data that the thresholding method exhibits attenuation bias and a bias-variance tradeoff as a function of the selected threshold, and we recover these phenomena in precise mathematical terms in an idealized Gaussian model. Second, we introduce a new method for estimation and inference in single-cell CRISPR screens that accounts for the measurement error inherent in the experiment. The method, called \textit{GLM-EIV} (generalized linear model with errors-in-variables), implicitly estimates the probability that each cell was perturbed, obviating the need to explicitly impute perturbation assignments via thresholding or another heuristic. Theoretical analyses and simulation studies indicate that GLM-EIV outperforms the thresholding method in large regions of the parameter space.

@@

\begin{equation}\label{theoretical_model}
\begin{cases}
m_i = \beta^m_0 + \beta^m_1 p_i + \ep_i \\
g_i = \beta^g_0 + \beta^g_1 p_i + \tau_i \\
p_i \sim \textrm{Bern}(\pi) \\
\ep_i, \tau_i \sim N(0,1) \\
p_i \indep \tau_i \indep \ep_i.
\end{cases}
\end{equation}
For a given threshold $c \in \R$, let $\hat{p}_i = \mathbb{I}(g_i \geq c).$ The thresholding estimator $\hat{\beta}^m_1$ for $\beta^m_1$ is given by $$\hat{\beta}^m_1 = \frac{\sum_{i=1}^n (\hat{p}_i - \overline{\hat{p}}) (m_i - \overline{m})}{\sum_{i=1}^n (\hat{p}_i - \overline{\hat{p}})}.$$

\begin{proposition}
 The limit in probability (as $n \to \infty$) of $\hat{\beta}^m_0$ is
\begin{equation}\label{thresh_est_intercepts}
\hat{\beta}^m_1 \xrightarrow{P} \beta^m_1 \left(\frac{ \pi( \omega - \E[ \hat{p}_i ])}{ \E[\hat{p}_i] (1 - \E[\hat{p}_i])}\right),
\end{equation}
where
$$\begin{cases}
\omega = 1 - \Phi\left(c - \beta_1^g - \beta_0^g \right) ,\\ \zeta = 1 - \Phi\left( c - \beta^g_0 \right), \\ \E[\hat{p}_i] = \zeta(1-\pi) + \omega\pi. 
\end{cases}$$
\end{proposition}
Let $\gamma: \R^4 \to \R$ be defined by
$$ \gamma( \beta^g_0, \beta^g_1, c, \pi ) = \frac{\pi (\omega - \E[\hat{p}_i])}{\E[\hat{p}_i] (1 - \E[\hat{p}_i]) }.$$ We call $\gamma$ the ``attenuation function.'' Observe that \begin{itemize}
\item[i.] $\gamma$ does not depend on $\beta^m_1$ or $\beta^m_0$, and
\item[ii.] $\hat{\beta}^m_1 \xrightarrow{P} [\gamma(\beta_0^g, \beta_1^g, c, \pi)] \beta^m_1.$
\end{itemize}
Let $b: \R^4 \to \R$ be the asymptotic bias of $\hat{\beta}^m_1$. We can express $b$ in terms of $\gamma$ as $$b\left(\beta^g_0, \beta^g_1, c, \pi \right) = \beta^m_1 - \gamma( \beta_0^g, \beta_1^g, c, \pi) \beta^m_1 = \beta^m_1 \left[ 1- \gamma(\beta^g_0, \beta^g_1, c, \pi) \right].$$ The asymptotic bias vanishes when the attenuation function equals $1$.

@@

\begin{itemize}
\item \begin{proposition}
The Bayes-optimal threshold $c_\textrm{bayes}$ is a critical value of the limiting bias $b$ (holding other parameters fixed; Figure \ref{thresholding_theoretical}a, all panels).
\end{proposition}
For certain values of $\beta^g_0$ and $\beta^g_1$, $c_\textrm{bayes}$ is a maximizer of the bias (Figure \ref{thresholding_theoretical}a, left); for other values of $\beta^g_0$ and $\beta^g_1$, $c_\textrm{bayes}$ is a minimizer of the bias (Figure \ref{thresholding_theoretical}a, right).
\item
\begin{proposition}
The limit of the bias in $c$ is $1/2$, i.e. 
$$\lim_{c \to \infty} b(\beta^g_0, \beta^g_1, c, 1/2) = 1/2.$$
\end{proposition}
 In other words, we always can set the threshold to a large number and achieve a bias of $1/2$ (Figure \ref{thresholding_theoretical}a, all panels). This fact establishes an upper bound on the bias of thresholded regression (under optimal threshold selection).
 \item
 \begin{proposition}
 For $\beta^g_1 < 2 \Phi(3/4) \approx 1.35$, $c = c_\textrm{bayes}$ yields a \textit{larger} bias than $c = \infty$ (Figure \ref{thresholding_theoretical}a, left); for $\beta^g_1 > 2 \Phi(3/4)$, $c_\textrm{bayes}$ yields a \textit{smaller} bias than $c = \infty$ (Figure \ref{thresholding_theoretical}a, right); finally, for $\beta^g_1 = 2 \Phi(3/4)$, $c = c_\textrm{bayes}$ and $c = \infty$ both yield a bias of $1/2$ (Figure \ref{thresholding_theoretical}a, middle).
 \end{proposition}
\item \begin{proposition}
Limiting bias decreases monotonically as $\beta_1^g$ increases (Figure \ref{thresholding_theoretical}a; fix threshold and scan panels from left to right).
\end{proposition}
This is consistent with the intuition that the problem becomes easier as the gRNA mixture distribution becomes increasingly well-separated.
\item 
\begin{proposition} For all $\beta^g_0, \beta^g_1,$ and $c \in \R,$ the bias is strictly greater than $0$ and less than $2$ (Figure \ref{thresholding_theoretical}a, all panels). 
\end{proposition}
Therefore, the thresholding method suffers from attenuation bias.
\end{itemize}

@@

Next, we generalize the expression for bias when the threshold is large to arbitrary $\pi \in [0,1/2]$:
\begin{proposition}
The following relationship holds:
$$ \lim_{ c \to \infty } b(\beta^g_0, \beta^g_1, c, \pi) = \pi.$$
\end{proposition}
In other words, when the perturbation probability is equal to $\pi$ and the threshold large, the bias is $\pi$. We can understand this result intuitively by considering an extreme example. When $\pi$ is very small (e.g., $\pi = 0.01$), almost all cells are unperturbed. Therefore, if we select a large threshold, we correctly classify nearly all unperturbed cells as unperturbed; on the other hand, the \textit{perturbed} cells that we misclassify as \textit{unperturbed} are swamped in number by the truly unperturbed cells, resulting in a small bias.

@@

% Finally, to shed light on limitations of the large threshold selection strategy, we study the variance of the thresholding estimator. Let $v$ denote the limiting variance of scaled estimator $\sqrt{n} \hat{\beta}^m_1$, i.e.,
% $$v(\beta^g_0, \beta^g_1, c, \pi) = \lim_{n \to \infty} \V(\sqrt{n} \hat{\beta}^m_1).$$ The limiting variance of $\sqrt{n} \hat{\beta}^m_1$ grows without bound as the threshold $c$ tends to infinity: 
% $$ \lim_{c \to \infty} v(\beta^g_0, \beta^g_1, c, \pi) = \infty.$$ Thus, while selecting a large threshold is effective from a bias reduction standpoint when $\pi$ is small, it causes the variance to explode.

To more precisely characterize the bias-variance tradeoff of the thresholding method, we consider a slightly simpler model: we set $\beta^m_0$ and $\beta^g_0$ to zero in (\ref{theoretical_model}) and study the no-intercept least squares estimator
$$ \hat{\beta}^m_1 =  \frac{ \sum_{i=1}^n \hat{p}_i m_i }{ \sum_{i=1}^n \hat{p}_i^2}.$$ We uncover an exact bias-variance tradeoff as a function of the selected threshold (Figure \ref{thresholding_theoretical}c).

\begin{proposition}
The limiting distribution is:
$$\sqrt{n}(\hat{\beta}^m - \beta^m) \xrightarrow{d} N(0, v).$$
\end{proposition}

The bias-variance analysis reveals that the best strategy for minimizing mean-squared error is to select a threshold that induces moderate bias. (A downside of this approach, of course, is that constructing valid confidence intervals becomes more challenging.)

@@

\subsubsection{Bayes-optimal decision boundary as a critical value of $\gamma$}
Let $c_\textrm{bayes} = \beta^g_0 + (1/2)\beta^g_1.$ We show that $c = c_\textrm{bayes}$ is a critical value of $\gamma$ for $\pi = 1/2$ and given $\beta^g_1$, i.e, $$D_2 \gamma (\beta^g_1, c_\textrm{bayes}, 1/2) = 0.$$ By the quotient rule,
\begin{multline}\label{quotient_rule}
D_2\gamma(\beta^g_1, c, 1/2) \\ = \frac{(1/2) D_2g(\beta^g_1, c, 1/2) h(\beta^g_1, c, 1/2) - (1/2) g(\beta^g_1, c, 1/2) D_2 h(\beta^g_1, c, 1/2)}{[h(\beta^g_1, c, \pi)]^2} \\ = \frac{\left[4(1/2) D_2g(\beta^g_1, c, 1/2)\right] h(\beta^g_1, c, 1/2) - (1/2) g(\beta^g_1, c, 1/2) \left[4 D_2 h(\beta^g_1, c, 1/2)\right]}{[2h(\beta^g_1, c, \pi)]^2},
\end{multline}
where the second equality multiplies both numerator and denominator by $4$. We have by (\ref{dg_dc}) that
\begin{multline}\label{dg_dc_bayes}
4 \cdot (1/2) \cdot D_2 g(\beta^g_1, c_\textrm{bayes}, 1/2)  = f( \beta^g_1/2) - f( -\beta^g_1/2) \\ = f(\beta^g_1/2) - f(\beta^g_1/2) = 0.
\end{multline}
Next, we have by (\ref{dh_dc}) that
\begin{multline}\label{dh_dc_bayes}
4 \cdot D_2 h(\beta^g_1, c_\textrm{bayes}, \pi) = [f( \beta^g_1/2) + f( -\beta^g_1/2)] \cdot \\ \left[  \int_{-\infty}^{-\beta^g_1/2} f + \int_{-\infty}^{\beta^g_1/2} f - \int_{-\infty}^{\beta^g_1/2} f - \int_{-\infty}^{-\beta^g_1/2} f \right] = 0.
\end{multline}
Plugging in (\ref{dh_dc_bayes}) and (\ref{dg_dc_bayes}) to (\ref{quotient_rule}), we find that 
$$D_2(\beta^g_1, c_\textrm{bayes}, 1/2) = 0.$$ It follows that $c_\textrm{bayes}$ is a critical value of the bias function $b$.

\subsubsection{Comparison of Bayes-optimal decision boundary to large threshold}

We study the bias produced by setting the threshold to a large number versus setting the threshold to the Bayes-optimal decision boundary. Let $r: \R \to \R$ be the value of attenuation function evaluated at the Bayes-optimal decision boundary $c_\textrm{bayes} = \beta^g_0 + (1/2) \beta^g_1$:
\begin{multline*}
r(\beta^g_1) = \gamma(\beta^g_1, \beta^g_0 + (1/2)\beta^g_1, 1/2) = \frac{\Phi(\beta^g_1/2) - \Phi(-\beta^g_1/2)}{\left[\Phi(-\beta^g_1/2) + \Phi( \beta^g_1/2) \right] \left[\Phi(\beta^g_1/2) + \Phi( -\beta^g_1/2)\right]} \\ = \frac{\int_{-\beta^g_1/2}^{\beta^g_1/2} f}{\left[ 1 - \Phi(\beta^g_1/2) + \Phi(\beta^g_1/2) \right]\left[ \Phi(\beta^g_1/2) + 1 - \Phi(\beta^g_1/2) \right]} = 2 \int_{0}^{\beta^g_1/2} f = 2 \Phi(\beta^g_1/2) - 1.
\end{multline*}
We set $r$ to $1/2$ and solve for $\beta^g_1$:
$$ 2\Phi(\beta^g_1/2) -1 = 1/2 \iff \Phi(\beta^g_1/2) = 3/4 \iff \beta^g_1 = 2 \Phi^{-1}(3/4) \approx 1.35.$$ The bias induced selecting a large threshold is $1/2$, i.e.
$$\gamma(\beta^g_1, \infty, 1/2) = 1/2.$$
Additionally, $r$ is a strictly increasing function. Therefore, $\gamma(\beta^g_1, c_\textrm{bayes},1/2) < \gamma(\beta^g_1, \infty, 1/2)$ on $\beta^g_1 \in [0, 2\Phi^{-1}(3/4))$; $\gamma(\beta^g_1, c_\textrm{bayes},1/2) = \gamma(\beta^g_1, \infty, 1/2)$ for $\beta^g_1 = 2\Phi^{-1}(3/4)$; and $\gamma(\beta^g_1, c_\textrm{bayes},1/2) > \gamma(\beta^g_1, \infty, 1/2)$ on $\beta^g_1 \in (2\Phi^{-1}(3/4), \infty)$. Because $b(\beta^g_1, c, 1/2) = \beta^m_1(1 - \gamma(\beta^g_1, c, 1/2))$ similar results hold for the bias (but with the inequalities flipped).

\subsubsection{Limit in $\beta^g_1$}
We compute the limit of $\gamma$ in $\beta^g_1$ given $\pi = 1/2$.  First,
\begin{multline*}
 \lim_{\beta^g_1 \to \infty} \gamma(\beta^g_1, c, 1/2) = \frac{ 1 - \int_{-\infty}^{\beta^g_0 - c} f}{\left[ 1 + \int_{-\infty}^{ \beta^g_0 - c} f \right] \left[\int_{-\infty}^{ c - \beta^g_0} f \right]} \\ = \frac{\int_{-\infty}^{c - \beta^g_0} f }{ \left[1 + \int_{-\infty}^{ \beta^g_0 - c} f \right] \left[\int_{-\infty}^{ c - \beta^g_0} f \right]} = \frac{1}{1 + \int_{-\infty}^{\beta^g_0 - c} f} < 1.
\end{multline*}
Similarly,
\begin{multline*}
\lim_{\beta^g_1 \to -\infty} \gamma(\beta^g_1, c, 1/2) = \frac{ -\int_{-\infty}^{\beta^g_0 - c} f }{\left[ \int_{-\infty}^{\beta^g_0 - c} f \right] \left[ \int_{-\infty}^{c - \beta^g_0} f + 1 \right]} = \frac{-1}{1 + \int_{-\infty}^{c - \beta^g_0} f} > -1.
\end{multline*}

\subsubsection{Monotonicity in $\beta^g_1$}
We show that $\gamma$ is monotonically increasing in $\beta^g_1$ for $\pi = 1/2$ and given $(\beta^g_1, \beta^g_0)$. We begin by stating and proving two lemmas. The first lemma establishes an inequality that will serve as the basis for the proof.

\begin{lemma}
The following inequality holds: 
\begin{multline}\label{basic_ineq_cp}
\left[\Phi(\beta^g_0 - c) + \Phi(\beta^g_0 + \beta^g_1 - c) \right] \\ \cdot \left[\Phi(\beta_0^g + \beta_1^g - c) - \Phi(\beta_0^g - c) + \Phi(c - \beta^g_0) + \Phi(c - \beta_0^g - \beta_1^g) \right] \\ \geq \left[\Phi(\beta_0^g + \beta_1^g - c) - \Phi(\beta_0^g - c)\right]\left[\Phi(c - \beta^g_0) + \Phi(c - \beta_0^g - \beta_1^g)\right].
\end{multline}
\end{lemma}
\textbf{Proof}: We take cases on the sign on $\beta^g_1$.

\noindent
\underline{Case 1}: $\beta^1_g < 0$. Then $ \beta^g_1 + (\beta^g - c) < (\beta^g_0 - c),$ implying $\Phi(\beta^g_0 + \beta^g_1 - c) < \Phi(\beta^g_0 - c),$ or $[\Phi(\beta^g_0 + \beta^g_1 - c) - \Phi(\beta^g_0 - c)] < 0.$ Moreover, $[\Phi(c - \beta^g_0) + \Phi(c - \beta_0^g - \beta_1^g)]$ is positive. Therefore, the right-hand side of (\ref{basic_ineq_cp}) is negative.

Turning our attention of the left-hand side of (\ref{basic_ineq_cp}), we see that
\begin{equation}\label{basic_ineq_cp_2}
\Phi(\beta^g_0 + \beta^g_1 - c) + \Phi( c - \beta^g_0 - \beta^g_1) = 1 -\Phi(\beta^g_0 + \beta^g_1 - c) + \Phi( c - \beta^g_0 - \beta^g_1) = 1.
\end{equation}
Additionally, $\Phi(\beta^g_0 - c) < 1$ and $ \Phi(c - \beta^g_0) > 0$. Combining these facts with (\ref{basic_ineq_cp_2}), we find that
$$ \left[\Phi(\beta_0^g + \beta_1^g - c) - \Phi(\beta_0^g - c) + \Phi(c - \beta^g_0) + \Phi(c - \beta_0^g - \beta_1^g) \right] > 0. $$ Finally, because $\left[\Phi(\beta^g_0 - c) + \Phi(\beta^g_0 + \beta^g_1 - c) \right] > 0,$ the entire left-hand side of (\ref{basic_ineq_cp}) is positive.
